Near Optimal Behavior with Approximate Abstraction

### 2016-02-04 paper/background.tex
Made background optimal policy consistent with rest of paper

The objective of an agent is to solve for the policy that maximizes its expected discounted reward from any state, denoted $\pi^*$. We denote the expected discounted reward for following policy $\pi$ from state $s$ as the value of the state under that policy, $V^\pi(s)$. We similarly denote the expected discounted reward for taking action $a \in \mathcal{A}$ and then following policy $\pi$ from state $s$ forever after as $Q^\pi(s,a)$, defined by the Bellman Equation as:
\begin{equation}
Q^\pi(s,a) = \mathcal{R}(s,a) + \gamma \sum_{s'} \mathcal{T}(s,a,s') Q^\pi(s',\pi(s))
\end{equation}
Similarly, the value function defined under a given policy, denoted $V^\pi(s)$, is defined as:
\begin{equation}
V^\pi(s) = \max_a Q^\pi(s,a)
\end{equation}
Lastly, we denote the value and $Q$ functions under the optimal policy as $V^*$ and $Q^*$, respectively. For further background, see~\cite{kaelbling1996reinforcement}.

@@

% Subsection: Exact Abstraction Framework
\subsection{Exact Abstraction Framework}

\citeauthor*{li2006towards} developed a framework for state abstraction in \acp{MDP}, the notation for which is defined in the previous section. Here, we survey their main results, which we then extend to the framework of approximate abstraction.

In particular, the authors define the following five classes of state-aggregation functions, inspired by existing methods for state abstraction in \acp{MDP}. For brevity, we include those that we build upon directly. (See their paper for full details.)

\bdefn{$\phi_{Q^*}$}
A $Q^*$-irrelevance abstraction, $\phi_{Q^*}$, is one such that:
\begin{equation}
\phi_{Q^*}(s_1) = \phi_{Q^*}(s_2) \rightarrow \forall a\ \left(Q^*(s_1,a) = Q^*(s_2,a)\right).
\end{equation}
\edefn

\bdefn{$\phi_{model}$}
A model-irrelevance abstraction, $\phi_{model}$, is one such that:
\begin{multline}
\phi_{model}(s_1) = \phi_{model}(s_2) \rightarrow \forall a\ \mathcal{R}_G(s_1,a) = \mathcal{R}_G(s_2,a).
\end{multline}
and
\begin{multline}
\phi_{model}(s_1) = \phi_{model}(s_2) \rightarrow \\ \forall a \sum_{s' \in \phi^{-1}(\phi(s_1))} \mathcal{T}_G(s_1,a,s') = \sum_{s' \in \phi^{-1}(\phi(s_2))} \mathcal{T}_G(s_2,a,s').
\end{multline}
\edefn

Intuitively, a $Q^*$-irrelevance is one that is guaranteed to compress states that share optimal Q functions. Critically, the direction of the implication is such that the abstraction is not guaranteed to abstract all states for which this condition is true, only that, if the abstraction function collapses two (or more) states to an abstract state, that those states have the same $Q^*$. Computing such a function is as difficult as solving the MDP, as it requires knowledge of the $Q*$, which is a sufficient condition for knowing the optimal policy. This is one of the motivations for investigating approximate abstractions, as we anticipate that an approximation of $Q^*$ is the sort of knowledge a learner can infer online.

A model-irrelevance abstraction is one that is guaranteed to compress states that transition to the same compressed states, and have identical rewards. Critically, this need not require that the ground transition functions of any two compressed states are identical; only that their corresponding abstract state transitions {\it in the abstract} MDP are identical. Unlike the $Q^*$ irrelevance, in the planning setting we anticipate knowing the full transition function and reward function up front, so it is feasible that an abstraction function of this kind may be used. In a model based learning setting, we expect an approximate abstraction akin to model-irrelevance to be more realistic, as we anticipate reinforcement learners to be able to approximate the information content of the transition function and reward function during learning.

>>>>

The objective of an agent is to solve for the policy that maximizes its expected discounted reward from any state, denoted $\pi^*$. We denote the expected discounted reward for following policy $\pi$ from state $s$ as the value of the state under that policy, $V^\pi(s)$. We similarly denote the expected discounted reward for taking action $a \in \mathcal{A}$ and then following policy $\pi$ from state $s$ forever after as $Q^\pi(s,a)$, defined by the Bellman Equation as:
\begin{equation}
Q^\pi(s,a) = \mathcal{R}(s,a) + \gamma \sum_{s'} \mathcal{T}(s,a,s') Q^\pi(s',\pi(s))
\end{equation}
Similarly, the value function defined under a given policy, denoted $V^\pi(s)$, is defined as:
\begin{equation}
V^\pi(s) = \max_a Q^\pi(s,a)
\end{equation}
Lastly, we denote the value and $Q$ functions under the optimal policy as $V^*$ or $V^{\pi^*}$ and $Q^*$ or $Q^{\pi^*}$, respectively. For further background, see~\cite{kaelbling1996reinforcement}.

@@

% Subsection: Exact Abstraction Framework
